---
title: "Reproducible Research: Peer Assessment 1"
output: 
html_document:
keep_md: true
---

```{r echo = TRUE}
# load all packages used in this exploratory analysis
library(knitr)
library(scales)
library(ggplot2)
library(xtable)

```

## Loading and preprocessing the data

```{r echo = TRUE}
# load data
data <- read.csv("activity.csv")

# Transform date attribute from Factor to Date 
data$date <- as.Date(data$date)

```

## What is mean total number of steps taken per day?

1.- Calculate the total number of steps taken per day

2.- If you do not understand the difference between a histogram and a barplot, research the difference between them. Make a histogram of the total number of steps taken each day

3.- Calculate and report the mean and median of the total number of steps taken per day

```{r echo = TRUE} 
# remove NA in data
data2 <- data[!is.na(data$steps),]

# Calculate the total number of steps taken per day
steps_by_day <- tapply(data2$steps, data2$date,  sum)
steps_by_day

# Make a histogram of the total number of steps taken each day
hist(steps_by_day, main = "Histogram of the total number of steps taken each day", xlab = "Steps by day")

```

The mean of the total number of steps taken per day is
```{r echo=TRUE, results="asis"}
mean(steps_by_day, na.rm = TRUE)
```

The median of the total number of steps taken per day is
```{r echo=TRUE, results="asis"}
median(steps_by_day, na.rm = TRUE)
```

## What is the average daily activity pattern?

Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r}
steps_by_interval <- tapply(data2$steps, data2$interval,  mean)

 plot(row.names(steps_by_interval), steps_by_interval, type = 'l', main = "Average number of steps taken by interval", xlab = "5-minute interval", ylab = "Averaged across all days")
```

Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
names(which.max(steps_by_interval))

```

## Imputing missing values

Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)

```{r echo=TRUE}
number_missing_values <- sum(is.na(data$steps))
```

The total number of rows with NAs is `r number_missing_values`

Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r}
data_fill <- data
 #The data is filled by mean value
data_fill[is.na(data_fill$steps),]$steps <- mean(data$steps, na.rm = TRUE)

```

Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. 
```{r}
# Calculate the total number of steps taken per day
steps_by_day <- tapply(data_fill$steps, data_fill$date,  sum)

# Make a histogram of the total number of steps taken each day
hist(steps_by_day, main = "Histogram of the total number of steps taken each day - NA repleced", xlab = "Steps by day")
```

The mean of the total number of steps taken per day is
```{r echo=TRUE, results="asis"}
mean(steps_by_day, na.rm = TRUE)
```

The median of the total number of steps taken per day is
```{r echo=TRUE, results="asis"}
median(steps_by_day, na.rm = TRUE)
```

Do these values differ from the estimates from the first part of the assignment? 
No, the values are not very differents.

What is the impact of imputing missing data on the estimates of the total daily number of steps?
Because I used the mean of values the data don't have many variation, may be with another method the changes can be more serious. 

## Are there differences in activity patterns between weekdays and weekends?

Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.

```{r}
week_days <- c('lunes', 'martes', 'miércoles', 'jueves', 'viernes')
data2$week_day <- factor((weekdays(data2$date) %in% week_days), levels=c(FALSE, TRUE), labels=c('weekend', 'weekday'))
summary(data2)
```

Make a panel plot containing a time series plot 

```{r}
qplot(interval, steps, data=data2, stat="summary", fun.y="mean", geom="line", ) + facet_wrap( ~ week_day, nrow = 2, ncol = 1) 

```
